Automatic Parallelization and Transparent Fault Tolerance

### 2016-07-12 pubs/tfp16-draft/paper.tex
Add GHC-Bridge section to TFP paper

>>>>

\section{(Ab)Using GHC}

The bridge to full Haskell is in progress, with GHC providing much of
the structure.  We are using GHC as a library (as opposed to modifying
its source to use our code generation and runtime system).  This
choice was made primarily due to the size and complexity of a project
as large as GHC.  Fully understanding and modifying even its build
system would be a significant undertaking.

We are using GHC's pipeline to translate Haskell source to GHC's
representation of the STG language, which we then translate to our
own.  Up to the STG generation stage, this usage is equivalent to
GHC's \texttt{--make} behavior, which allows us to benefit from the
module dependency analysis, typechecking, Core-level transformations
and informative error messages for invalid programs.

Though we follow the same rough idea as GHC in implementing the STG
abstract machine, this bridging process is not entirely
straightforward.  The challenges we've seen fit into two categories.

\subsection{Producing Full STG Programs}

When compiling a standard Haskell program with GHC, modules imported
from other packages have already been compiled, and GHC simply links
the user object code with them.  Our compiler has no facility for
linking GHC-produced object code, so we need it to produce
\textit{full} STG programs built up from primitive operations.  GHC
does not offer object-to-STG translation, so this means we must
produce our own Prelude and supporting Base libraries to be included
with every program, and decoupling Haskell from GHC's wired-in types
as much as possible.

This is partially accomplished with extensions like
\texttt{NoImplicitPrelude} and \texttt{RebindableSyntax} (which
implies the former).  We cannot get a complete separation, however:
Numeric literals are tied to GHC's \texttt{Integer} and
\texttt{Rational} types, and guards are tied to GHC's \texttt{Bool}.
Lists and tuples, being special syntactic forms are also tightly
wired-in. To get around this, we provide placeholder definitions that
parallel any that we cannot replace, and use them to crudely translate
the exceptional cases.

\subsection{Navigating GHC's Internals}

One of the goals of using GHC as a library was to avoid being
overwhelmed with the necessary complexity of such a large and mature
tool, but this has not been entirely unavoidable.  Simply producing
STG without incurring any (non-temporary) file creation proved not to
be a difficult task, and can be accomplished using only the interface
exposed via the primary API, the \texttt{GHC} module.  This abstract
STG is even ostensibly very close to our own, but the auxiliary data
structures (e.g. those encapsulating identifier information and types)
are rich and deeply nested.  Determining what is strictly necessary
for our backend and what we have no implementation of (e.g. the
Foreign Function Interface) has taken considerable time and effort.

It is a testament to the talents of the GHC contributors and
maintainers that we have managed to come even this far.  Their
documentation and exposed interfaces have made navigating and
comprehending the project internals much more pleasant than expected.
